\chapter{Problem Analysis and Requirements}

This chapter contains an analysis of the problem and breaks it down into sub problems. It then sets out the functional and non-functional requirements.

\section{Problem Analysis}
The problem can be broken down into three main components:

\subsection{Video Game Development}
Creating a basic video game involves designing core gameplay mechanics, implementing player controls, creating a game environment, and establishing basic game rules and objectives. 
This requires selecting an appropriate game engine that supports the planned features and performance requirements. 
Core gameplay mechanics and rules must be defined to provide an engaging player experience, while visual assets and user interface elements need implementation to create a cohesive visual identity. 
Player control systems need to be intuitive and responsive, allowing for meaningful interaction with the game world. Finally, clear win/lose conditions need establishment to provide players with goals and motivation.

\subsection{Network Communication Interface}
Connecting an external AI agent to the game requires establishing a reliable communication channel. 
This involves designing a networking protocol that efficiently transmits necessary game state information and receives AI decisions. 
A server-client architecture must be implemented to facilitate this exchange, along with data serialization and parsing mechanisms to ensure proper interpretation of messages on both ends. 
Synchronization between game state and AI agent is critical to maintain consistency, while also managing network latency and potential disconnections to ensure smooth gameplay even under non-ideal network conditions.

\subsection{AI Integration into Gameplay}
Incorporating the AI agent as part of the gameplay loop involves defining the AI's role within the game context, whether as an opponent, ally, or environmental element. 
Interfaces for AI action and observation must be created to allow the AI to both perceive the game state and affect it meaningfully. 
Game state representation needs careful implementation to provide the AI with relevant information in a consumable format. The AI's capabilities must be balanced with player experience to ensure engaging but fair gameplay. 
Additionally, feedback mechanisms must be designed to showcase AI behavior, allowing players to understand and respond to AI actions appropriately.

\section{Functional Requirements}
\begin{table}[H]
    \centering
    \begin{tabular}{|p{0.15\textwidth}|p{0.7\textwidth}|}
        \hline
        \textbf{ID} & \textbf{Requirement} \\
        \hline
        FR1 & The system must implement a playable video game with clear objectives and mechanics \\
        \hline
        FR2 & The game must provide intuitive player controls and interface \\
        \hline
        FR3 & The system must implement a client-server network architecture for AI communication \\
        \hline
        FR4 & The system must be able to transmit game state information to an external AI agent \\
        \hline
        FR5 & The system must be able to receive and process action commands from an external AI agent \\
        \hline
        FR6 & The game must incorporate AI actions into the gameplay loop in real-time \\
        \hline
        FR7 & The system must provide a complete representation of relevant game state to the AI \\
        \hline
        FR8 & The game must provide visual feedback of AI actions to the human player \\
        \hline
        FR9 & The system must handle connection initialization and termination with the AI agent \\
        \hline
        FR10 & The game must implement win/lose conditions that account for AI participation \\
        \hline
    \end{tabular}
    \caption{Functional Requirements}
\end{table}

\section{Non-functional Requirements}
\begin{table}[H]
    \centering
    \begin{tabular}{|p{0.15\textwidth}|p{0.7\textwidth}|}
        \hline
        \textbf{ID} & \textbf{Requirement} \\
        \hline
        NFR1 & The game must maintain a minimum framerate of 30 FPS during gameplay \\
        \hline
        NFR2 & Network communication latency should not exceed 100ms under normal conditions \\
        \hline
        NFR3 & The system must recover gracefully from network disconnections \\
        \hline
        NFR4 & The game should be accessible to users with minimal gaming experience \\
        \hline
        NFR5 & The communication protocol must be well-documented for AI integration \\
        \hline
        NFR6 & The game's codebase must be maintainable and well-structured \\
        \hline
        NFR7 & The system should provide logging capabilities for debugging AI interactions \\
        \hline
        NFR8 & The game must run on standard desktop hardware without specialized requirements \\
        \hline
    \end{tabular}
    \caption{Non-functional Requirements}
\end{table}